Implementing simplified INLA into `aghq`

### 2022-11-22 src/dev_sinla/sinla.Rmd
logSumExp, simplify writing in Gaussian AGHQ, use AGHQ to pick x[i] nodes, use spline interpolation for plotting marginal

We're going to start by normalising $\tilde p_\text{LA}(\theta, y)$.
Remember that we did this above in the empirical Bayes case, but found that it wasn't strictly necessary.

@@

If you wanted to evaluate this to obtain each marginal then it would be something like (pseudo-code):

```{r eval=FALSE}
for(i in 1:n) {
  for(x in x_vals) {
    for(z in z_vals) {
      #' ...
    }
  }
}
```

To start with, let's evaluate this for `i = 1`, `x = x_test_value` and the first value of `z` (actually I don't really know what $z$ runs over, but this is just the first weight):

```{r}
x_test_value <- 0
z <- 1
i <- 1
```

The weight $\omega(z)$ is:

The normalised Laplace approximation $\tilde p_\text{LA}(\theta(z) \, | \, y)$ is:

```{r}
theta <- as.numeric(modesandhessians[z, thetanames])
- h_norm(theta)
```

The normalised Gaussian approximation $p_\text{G}(x_i \, | \, \theta(z), y)$ on the log-scale is:

```{r}
mode_i <- as.numeric(modesandhessians$mode[[z]][i])
Sigma <- solve(modesandhessians$H[[z]])
sd_i <- as.numeric(sqrt(Sigma[i, i]))

dnorm(x_test_value, mean = mode_i, sd = sd_i, log = TRUE)
```

So, putting it together, the value we would compute on the log-scale is as follows:

```{r}
dnorm(x_test_value, mean = mode_i, sd = sd_i, log = TRUE) + (- h_norm(theta)) + log(modesandhessians$weights[z])
```

To obtain the unlogged version we can use:
$$
\exp \left( \log\tilde p_\text{G}(x_i \, | \, \theta(z), y) + \log \tilde p_\text{LA}(\theta(z) \, | \, y) + \log \omega(z) \right)
$$

```{r}
exp(dnorm(x_test_value, mean = mode_i, sd = sd_i, log = TRUE) + (- h_norm(theta)) + log(modesandhessians$weights[z]))
```

Great -- now let's sum over `z` so that we can obtain the probability for our `x_test_value` in marginal `i`.
We won't worry about inefficient R for loops for the time being.

for(z in 1:nrow(modesandhessians)) {
  theta <- as.numeric(modesandhessians[z, thetanames])
  mode_i <- as.numeric(modesandhessians$mode[[z]][i])
  Sigma <- solve(modesandhessians$H[[z]])
  sd_i <- as.numeric(sqrt(Sigma[i, i]))
  pdf <- pdf + exp(dnorm(x_test_value, mean = mode_i, sd = sd_i, log = TRUE) + (- h_norm(theta)) + log(modesandhessians$weights[z]))
}

Using a `fine_grid` set of values for `x_vals` for each `i` might be overly computationally expensive.
Instead what we can do is perform another separate AGHQ to select a set of `x_vals`.
To do this with the `aghq` R package, we could create a `TMB` template to evaluate $\tilde p(x_i \, | \, y)$, but I imagine this might be too annoying.
Instead, we could create the `ff` object (with components `ff$fn`, `ff$gn` `ff$hn`) in R directly.

>>>>

We're going to start by normalising $\tilde p_\text{LA}(\theta, y)$.
We did this above in the empirical Bayes case, but found that it wasn't strictly necessary.

@@

To start with, let's evaluate this on the log-scale for `i = 1`, `x = x_test_value` and the first value of `z`.
The weight $\omega(z)$ will be `modesandhessians$weights[z]`.
The normalised Laplace approximation $\tilde p_\text{LA}(\theta(z) \, | \, y)$ will be `- h_norm(as.numeric(modesandhessians[z, thetanames]))`.
And to obtain the normalised Gaussian approximation $p_\text{G}(x_i \, | \, \theta(z), y)$ on the log-scale we get the relevant subcomponents of the joint Gaussian then use `ndorm` with `log = TRUE`.

```{r}
x_test_value <- 0

#' Note that in some sense z actually runs over nodes, not indices
z <- 1 
i <- 1

mode_i <- as.numeric(modesandhessians$mode[[z]][i])
Sigma <- solve(modesandhessians$H[[z]])
sd_i <- as.numeric(sqrt(Sigma[i, i]))

dnorm(x_test_value, mean = mode_i, sd = sd_i, log = TRUE) + (- h_norm(theta)) + log(modesandhessians$weights[z])
```

Now let's sum over `z` so that we can obtain the log-probability for our `x_test_value` in marginal `i`:

```{r}
gaussian_marginal <- function(x) {
  lp <- vector(mode = "numeric", length = nrow(modesandhessians))

  for(z in 1:nrow(modesandhessians)) {
    theta <- as.numeric(modesandhessians[z, thetanames])
    mode_i <- as.numeric(modesandhessians$mode[[z]][i])
    Sigma <- solve(modesandhessians$H[[z]])
    sd_i <- as.numeric(sqrt(Sigma[i, i]))
    lp[z] <- dnorm(x, mean = mode_i, sd = sd_i, log = TRUE) + (- h_norm(theta)) + log(modesandhessians$weights[z])
  }

  matrixStats::logSumExp(lp)
}

gaussian_marginal(x_test_value)
```

To pick a set of $x_i$ to evaluate the marginal at, we will use an approximation to the AGHQ nodes.
We will take the hyperparameters to be fixed at their mode, then choose the appropriate marginals of the joint Gaussian:

```{r}
#' Get the row number of modesandhessians which contains that theta mode
theta_mode_location <- which.max(quad$normalized_posterior$nodesandweights$logpost_normalized)

#' These are the mode and standard deviation from the Gaussian approximation at the node which is the mode of the Laplace approximation
#' (Say that three times fast...)
mode_i <- modesandhessians[[theta_mode_location, "mode"]][i]
sd_i <- sqrt(1 / modesandhessians[[theta_mode_location, "H"]][i, i])

#' Create Gauss-Hermite quadrature
gg <- mvQuad::createNIGrid(dim = 1, type = "GHe", level = 7)

#' Adapt to mode_i and sd_i
mvQuad::rescale(gg, m = mode_i, C = sd_i^2)

#' These are our set of x input values
gg$nodes

#' The log-probabilities at the set of input values
lps <- sapply(gg$nodes, gaussian_marginal)

#' Build a Lagrange polynomial interpolant of the marginal posterior
ss <- splines::interpSpline(gg$nodes, lps, bSpline = TRUE, sparse = TRUE)

interpolant <- function(x) {
  as.numeric(stats::predict(ss, x)$y)
}

finegrid <- seq(-5, 5, by = 0.1)

df <- data.frame(x = finegrid, y = exp(interpolant(finegrid)))

ggplot(df, aes(x = x, y = y)) +
  geom_line() +
  theme_minimal() +
  labs(x = "phi_prev[1]", y = "Posterior")
```
